---
title: "Getting Started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r}
library(BASSr)
library(sf)
library(ggplot2)
library(dplyr)
```


## Setup

1. Get spatial data - Hexagonal grid (primary spatial units - PSU) with land cover characteristics for each hex
2. Get cost data
3. Run `full_BASS_run()`    


### 1. Spatial Data
- Spatial data frame (sf object) with 
    - Columns Cell/Hex ID (e.g., `ET_Index`, `hex_id`)
    - Columns defining land cover (e.g., CORINE land cover CLC `CLC15_1`)
- This must be either POINT or (MULTI)POLYGON (will be converted to points)

> Land cover characteristics should not be percentages, but should be XXXX?


**Clean up hex data**
```{r}
psu_hex <- clean_land_cover(psu_hex_dirty,  pattern = "CLC0013_") %>%
  units::drop_units() # for plotting, get rid of m^2

ggplot(data = psu_hex, aes(fill = LC01)) +
  geom_sf() +
  scale_fill_viridis_c()
```


### Basic Run
```{r}
d <- full_BASS_run(land_hex = psu_hex, 
                   num_runs = 10,
                   n_samples = 3)

ggplot(data = d, aes(colour = benefit)) +
  geom_sf() +
  labs(colour = "Benefit") +
  scale_colour_viridis_c()

d_hex <- left_join(psu_hex, st_drop_geometry(d), by = "hex_id")

ggplot(data = d_hex, aes(fill = benefit)) +
  geom_sf() +
  labs(fill = "Benefit") +
  scale_fill_viridis_c()
```

### Including Costs

```{r}
d <- full_BASS_run(land_hex = psu_hex, 
                   num_runs = 10,
                   n_samples = 3,
                   costs = psu_costs)

d_hex <- left_join(psu_hex, st_drop_geometry(d), by = "hex_id")

ggplot(data = d_hex, aes(fill = inclpr)) +
  geom_sf() +
  labs(fill = "Inclusion\nProbability") +
  scale_fill_viridis_c()
```

What if the costs of that highly beneficial points was much higher?

```{r}
which.max(d$benefit)
high_cost <- psu_costs
high_cost$RawCost[26] <- high_cost$RawCost[26] * 100

d <- full_BASS_run(land_hex = psu_hex, 
                   num_runs = 10,
                   n_samples = 3,
                   costs = high_cost)

d_hex <- left_join(psu_hex, st_drop_geometry(d), by = "hex_id")

ggplot(data = d_hex, aes(fill = inclpr)) +
  geom_sf() +
  labs(fill = "Inclusion\nProbability") +
  scale_fill_viridis_c()

```

Still high inclusion probability, but other points become relatively 'better'.

### Runs by 'hand'

```{r}
psu_hex <- clean_land_cover(psu_hex_dirty, pattern = "CLC0013_")
samples <- draw_random_samples(psu_hex, num_runs = 10, n_samples = 3)
benefit <- calculate_benefit(psu_hex, samples)
inc_prob <- calculate_inclusion_probs(benefit, costs = psu_costs)

d_hex <- left_join(psu_hex, st_drop_geometry(inc_prob), by = "hex_id")

ggplot(data = d_hex, aes(fill = inclpr)) +
  geom_sf() +
  labs(fill = "Inclusion\nProbability") +
  scale_fill_viridis_c()
```

Alternative pipe

```{r}
final <- psu_hex_dirty %>%
  clean_land_cover(pattern = "CLC0013_") %>%
  draw_random_samples(num_runs = 10, n_samples = 3) %>%
  calculate_benefit(psu_hex, samples = .) %>%
  calculate_inclusion_probs(costs = psu_costs)
```


### Weights? (e.g., Akimiski_Island.Rmd)


## Calculating Costs

- Need number of ARUs which to deploy

...


## Selection probabilities

### Simple selection

Here we'll sample 12 sites with a 20% over sample, resulting in a total of 14 
sites selected.

```{r}
g <- ggplot() + 
  geom_sf(data = psu_hex, fill = "white") +
  geom_sf(data = final, colour = "grey70")

g

sel <- run_grts_on_BASS(probs = final, 
                        num_runs = 1, 
                        nARUs = 12, 
                        os = 0.2)

sel_plot <- bind_rows(sel[["sites_base"]],
                      sel[["sites_over"]])

g + 
  geom_sf(data = sel_plot, aes(colour = siteuse), size = 5) +
  scale_colour_viridis_d(name = "Type of\nsites sampled", end = 0.7)

```

### Stratified selection

First let's create a dummy stratification and add it to our hexes for plotting
```{r}
final <- mutate(final, strat = c(rep("A", 15), rep("B", 18)))
psu_hex_strat <- select(final, "hex_id", "strat") |> 
  st_drop_geometry() |>
  left_join(psu_hex, y = _, by = "hex_id")

g <- ggplot() + 
  geom_sf(data = psu_hex_strat, aes(fill = strat), alpha = 0.4) +
  geom_sf(data = final, colour = "grey70")
g
```

Now we'll define how we want to sample these two strata.

Let's assume we don't really care about habitat A, so we don't want to sample
that one very much.

```{r}
nARUs <- list("A" = 2, "B" = 10)

sel <- run_grts_on_BASS(probs = final, 
                        num_runs = 1, 
                        stratum_id = strat,
                        nARUs = nARUs, 
                        os = 0.2)

sel_plot <- bind_rows(sel[["sites_base"]],
                      sel[["sites_over"]])

g + 
  geom_sf(data = sel_plot, aes(colour = siteuse), size = 5) +
  scale_colour_viridis_d(name = "Type of\nsites sampled", end = 0.7)

```

You can see that we've sampled much more of B than A, and that there are no 
over samples in A, which makes sense:

0.2 * 2 = 0.4 which rounds down to 0

If we wanted an over sample for A, we could define specific over sample amounts instead.

```{r}
nARUs <- list("A" = 2, "B" = 10)
os <- list("A" = 1, "B" = 4)

sel <- run_grts_on_BASS(probs = final, 
                        num_runs = 1, 
                        stratum_id = strat,
                        nARUs = nARUs, 
                        os = os,
                        seed = 123)

sel_plot <- bind_rows(sel[["sites_base"]],
                      sel[["sites_over"]])

g + 
  geom_sf(data = sel_plot, aes(colour = siteuse), size = 5) +
  scale_colour_viridis_d(name = "Type of\nsites sampled", end = 0.7)

```

Alternatively at this point (and especially with more strata) it might be easier
to supply a data frame rather than a series of lists.


```{r}
nARUs <- data.frame(n = c(2, 10),
                    strat = c("A", "B"),
                    n_os = c(1, 4))

sel <- run_grts_on_BASS(probs = final, 
                        num_runs = 1, 
                        stratum_id = strat,
                        nARUs = nARUs, 
                        seed = 123)

sel_plot <- bind_rows(sel[["sites_base"]],
                      sel[["sites_over"]])

g + 
  geom_sf(data = sel_plot, aes(colour = siteuse), size = 5) +
  scale_colour_viridis_d(name = "Type of\nsites sampled", end = 0.7)

```



