---
title: "Step by Step"
author: "David Hope"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    theme: united
highlight: github
vignette: >
  %\VignetteIndexEntry{Step by Step}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=T, warning=F, message =F}
knitr::opts_chunk$set(
  eval=T,
  collapse = TRUE,
  comment = "#>"
)
library(tidyverse)
library(sf)
library(spsurvey)
library(glue)
library(BASSr)
library(raster)
library(rlang)
library(patchwork)
ont.proj <- 3161

```


This is an example of how we calculate the benefit for one sample unit in a study area. The same approach applies to all other sample unit hexagons in the study area and also to calculating benefits for the study area hexagons in a region.


## Step 1 - Load in data

Here you will need a simple features (package sf) layer of hexagons for the study areas and sample units. Code to create the hexagons is available in `Hexagon Creation` vignette. The first step is to create or load both the study area and sample units. 

The data for the vignette is included in the package. It should load with the `BASSr` package. You'll need the objects `ontario`, `clrfile`, `lcc2015_codes`, `all_study_areas`, and `StudyArea_hexes`. Documentation for these objects is found by using the help function (eg `BASSr?ontario`). The raster for the example study area can be found in the system data (see chunk `raster-plot` below).

```{r load-data}


study_area_id  <- "ONT_SA_0740"

SA_sum <- StudyArea_hexes$landcover %>% as_tibble() %>% 
  summarize_at( .vars = vars(contains("LC")), .funs = sum) %>% 
   pivot_longer(cols = contains("LC"), names_to = "lc", values_to = "pHab_SA") %>% 
  mutate(pHab_SA = pHab_SA / sum(pHab_SA)) %>% 
  arrange(pHab_SA) %>% 
  mutate(cHAb_SA = cumsum(pHab_SA),
         lcFac = forcats::fct_reorder(lc, pHab_SA),
         lc_n = as.numeric(stringr::str_extract(lc, "\\d\\d") ))%>% 
  left_join(lcc2015_codes, by = c("lc_n" = "LCC_CODE")) %>% 
  mutate(lcc_fac = forcats::fct_reorder(LCC_NAME, pHab_SA))
```

```{r ontario-map, fig.cap='Map of Ontario study areas with example study area shown in red', fig.width=8, fig.height=8}
ggplot(all_study_areas) +
  geom_sf(data = ontario) +
  geom_sf(fill = 'white')+
  theme_linedraw() +
  geom_sf(data = filter(all_study_areas, StudyAreaID == study_area_id), fill = 'red')
```



## Step 2 Calculate Habitat Composition 
This example study area is shown below. The habitat is shown from the National Land Cover Classification 2015. The study area is `r nrow(StudyArea_hexes$landcover)` sample unit hexagons of `r round(mean(st_area(StudyArea_hexes$landcover)),0)`$m^2$.

```{r raster-plot, fig.cap = "Fig 1. The distribution of land cover classes across an example study area"}
r <- raster::raster(system.file("extdata", glue("{study_area_id}.tif"),
  package = "BASSr", mustWork = T)) %>% 
   
  raster::as.data.frame(xy=T, long =T) %>% as_tibble() %>%
    filter(!is.na(value)) %>% 
  left_join(lcc2015_codes, by = c("value" = "LCC_CODE"))


  pal_ <- clrfile$rgb
  names(pal_) <- as.character(clrfile$LCC_NAME)


  lc <- ggplot() + geom_raster(data = r, aes(x, y, fill = LCC_NAME)) +
    scale_fill_manual(values = pal_) +
    geom_sf(data = StudyArea_hexes$landcover, fill = NA) +
    labs(x = "", y = "", fill = "LCC2015") + theme_linedraw()
  
  # inplt <- ggplot() + geom_sf(data = StudyArea_hexes,
  #                             aes(fill = inclpr) ) + scale_fill_viridis_c()
  # library(patchwork)
  # lc + inplt

lc + ggspatial::annotation_scale(location = "bl", width_hint = 0.4) 
``` 


There is variation in the composition of land cover classes amongst sample units If we examine the increasing cumulative proportion of a sample unit hexagon covered by a given land cover type, we see a lot of variation in this distribtion between the `r nrow(StudyArea_hexes$landcover)` hexagons. To explore how this translates into a benefit we will focus on a single hexagon (shown in red below).

```{r land-cover-hexes, fig.cap = "Fig 2. The cumulative proportion of land cover classes within sample hexagons (grey) and the study area (black). The red line shows the example hexagon used in the initial benefit calculation.", fig.width= 8, fig.height=12}
hex_LC <- 
StudyArea_hexes$landcover %>% 
  as_tibble() %>% dplyr::select(-geometry) %>% 
  pivot_longer(cols = contains("LC"), names_to = "lc", values_to = "pHab") %>% 
  left_join(SA_sum, by =c( "lc")) %>% arrange(SampleUnitID, pHab_SA) %>% 
  group_by(SampleUnitID) %>% 
  mutate(cpHab = cumsum(pHab/100)) %>% ungroup %>% 
  mutate(lc_n = as.numeric(stringr::str_extract(lc, "\\d\\d") ))
set.seed(2277)
exampleHex <- sample_n(StudyArea_hexes$landcover, 1)
exLC <- hex_LC %>% filter(SampleUnitID %in% exampleHex$SampleUnitID)

ggplot(hex_LC, aes(lcc_fac, cpHab, group = SampleUnitID)) + 
  geom_step(colour = 'grey', 
            alpha = 0.1) + 
  geom_step(data = exLC, 
            colour = 'red')+
  geom_step(data = SA_sum, 
            aes(y = cHAb_SA, group = 1), 
            colour = "black", size =2) + theme_minimal() +
  labs(x = "Land Cover Class (Increasing freq from left to right)", y = "Cumulative proportion of hexagon") +
  theme(axis.text.x = element_text(family = 'mono', 
                                   angle= 75, vjust = 1, hjust = 1))

```

----- 

## Step 3 Sample unit hexagons - benefit calculation

### A. - Select a sample unit to calculate benefit for

Lets focus on one hexagon for now. It is shown in the figure above in red and below it is the red-shaded hexagon. We will call it the *focal sample unit* for now. 

```{r, fig.cap="Fig 3. Example Hexagon"}
ggplot(exampleHex) + geom_sf(fill = 'red') +
  geom_sf(data = StudyArea_hexes$landcover, fill = NA) + theme_minimal()


```

The focal sample unit varies in many ways from the study area. The table below compares how the percentages of land cover types differs between the example hexagon and the study area. We see for this example the example sample unit has a large proportion of "Temperate or sub-polar shrubland" relative to the study area, but the percentage of "Sub-polar or polar grassland-lichen-moss" is much lower than the study area. These differences with the study area will form the backbone of the benefit calculation for this focal sample unit.


```{r example-hex-table, results='asis'}

 exLC %>% dplyr::select(lcc_fac,pHab, pHab_SA) %>% 
  mutate(pHab_SA = pHab_SA * 100) %>% 
  knitr::kable(col.names = c("Land Cover", "% Example Sample Unit", "% Example Study Area"), digits = 2)

```


### B. - Draw a hypothetical sample set from all sample units in study area

To examine the importance of the focal sample unit, we first draw 1 random, spatially-dispersed sample set comprised of 10 sample unit hexagons. These 10 randomly drawn sample units are defined as the *hypothetical sample set*. They are used to estimate how the focal sample unit will increase (or decrease) the representivity of the sample set. 

```{r, fig.cap = "Fig 4. Plot of focal sample unit (red) and the hypothetical sample set (grey)"}

sample_hexes <- BASSr::draw_random_samples(att_cleaned = as_tibble(StudyArea_hexes$landcover), 
                                           att.sf = st_centroid(StudyArea_hexes$landcover), 
                                           num_runs = 1, nsamples = 10) 

sample_hexes_sf <- StudyArea_hexes$landcover %>% 
  filter(SampleUnitID %in% (sample_hexes$grts_random_sample %>% filter(panel == "PanelOne"))$SampleUnitID)

ggplot(exampleHex) + geom_sf(fill = 'red') +
  geom_sf(data = StudyArea_hexes$landcover, fill = NA) + 
  geom_sf(data = sample_hexes_sf, fill = 'grey')+
  theme_minimal()


```

The next step is to compare how the habitat representivity differs between the focal sample unit, the hypothetical sample set, the sample set with the sample unit included and the study area as a whole. Here we can see that the hypothetical sample set has a larger percentage of "Sub-polar taiga needleleaf forest" than the study area, while the focal sample unit has a lower percentage than the study area. When the focal sample unit is added to the sample set, the percentage shifts downwards and closer to the percentage for the study area. The opposite is true for "Temperate or sub-polar shrubland", where adding the focal sample unit moves the sample set plus focal sample unit away from the percentage observed at the study area. This is the fundamental measure of representivity for the focal sample unit. 


```{r}

sample_hexes_phab <-  hex_LC %>% filter(SampleUnitID %in% sample_hexes_sf$SampleUnitID) %>% 
  group_by(lcc_fac) %>% 
  summarize(pHab_samp = sum(pHab)) %>% ungroup %>% 
  mutate(pHab_samp = pHab_samp/ sum(pHab_samp)* 100)

m2toha <- 0.0001
study_area_size <- as.numeric(sum(st_area(StudyArea_hexes$landcover))) *m2toha
hexsize <- as.numeric(st_area(exampleHex))*m2toha


samp_com <- 
left_join( exLC %>% dplyr::select(lcc_fac,pHab, pHab_SA), sample_hexes_phab) %>% 
  mutate(pHab_Sample_plus_hex = 100*( (pHab*hexsize + pHab_samp*hexsize*10)/
                                        (sum(pHab*hexsize) + sum(pHab_samp*hexsize*10))),
         pHab_SA = pHab_SA * 100)

names(samp_com) <- c("lcFac", "Sample Unit", "Study Area", "Hypothetical Sample Set", "Sample Set + Sample Unit")

knitr::kable(samp_com, digits = 2)
```


## C. - Calculate representivity

As mentioned above, adding the focal sample unit to the sample set, shifts the habitat composition slightly and more so when the focal sample unit is quite different from the sample set. We now want to see if adding the sample unit to the hypothetical sample set will make the *updated* sample set more representative of the study area.


```{r, fig.cap = "Fig 5. Distribution of land cover by hexagon, sample, sample plus hexagon and study area", fig.width=14, fig.height=8}


d <- 
  samp_com %>%  
  pivot_longer(names_to = "sampletype", values_to = "phabitat", cols = -lcFac) %>% 
  mutate(sampletype = factor(sampletype, levels = c("Sample Unit", "Hypothetical Sample Set", "Sample Set + Sample Unit", "Study Area")))
                               # c("Hexagon",  "Sample", "Sample + Hex","Study Area")))


d%>% 
  ggplot(aes(sampletype, phabitat, fill = lcFac)) + geom_bar(stat= 'identity') +
  scale_fill_viridis_d() + labs(x = "" , y = "Percentage of total", fill = "Land Cover") + 
  cowplot::theme_cowplot()






```

-----



### D. - Calculate benefit for sample unit

To calculate the benefit for the focal hexagon, we determine for each land cover class if including the focal sample unit to the hypothetical sample set makes the sample set more or less representative. If, for a given land cover class, the hypothetical sample set becomes more similar to the study area, the benefit is the difference between the hypothetical sample set's representivity (here percentage of a land cover class) and the representivity if the sample unit is added into the sample set. Otherwise, the benefit for that land cover class is zero.

The benefit across all land cover types is summed to make the benefit value for that sample unit with that hypothetical sample set.



```{r, results = 'asis'}
rep_tabl <- samp_com %>% 
  # mutate_at(vars(-lcFac), ~(.*study_area_size))
  mutate(`Desired Direction` = case_when(`Study Area` > `Hypothetical Sample Set` ~ "Positive",
                                         `Study Area` < `Hypothetical Sample Set` ~ "Negative", 
                                         `Study Area` == `Hypothetical Sample Set` ~ "None"), 
         `Observed Direction` =  case_when(`Sample Set + Sample Unit` > `Hypothetical Sample Set` ~ "Positive",
                                         `Sample Set + Sample Unit` < `Hypothetical Sample Set` ~ "Negative", 
                                         `Sample Set + Sample Unit` == `Hypothetical Sample Set` ~ "None"),
         `Land Cover Benefit` = case_when(`Observed Direction` == "None"~"No Benefit",
           `Desired Direction`== `Observed Direction`~ "Benefit",
         TRUE~"No Benefit"), 
         Benefit = case_when(`Land Cover Benefit`== "Benefit"~abs(`Hypothetical Sample Set` - `Sample Set + Sample Unit`)/100,
                             TRUE~0) )


knitr::kable(rep_tabl, digits = 2)
```

In the table above, there are only 3 land cover types with a benefit greater than 0.005. As mentioned above "Temperate or sub-polar shrubland" moves the hypothetical sample set away from the study area percentage. However for "Sub-polar taiga needleleaf" the focal sample unit's low percentage is a benefit because it moves the hypothetical sample set percentage lower and closer to the study area's percent.


```{r, include =F, results = 'asis'}
a <- sum(rep_tabl$Benefit )
b <- quick_ben(d = exampleHex%>% as_tibble %>% 
             mutate_at(.vars = vars(contains("LC")),.funs = ~(.*hexsize)), 
          samples = sample_hexes$grts_random_sample %>% 
             summarize_at(.vars = vars(contains("LC")),.funs = ~sum(.*hexsize)), 
          land_cover_summary = SA_sum %>% mutate(ha = pHab_SA*study_area_size), 
          col_ = SampleUnitID, pd=F)

print(glue::glue("The benefit is {round(a,3)}. If it is calculated using the script it is {round(b$benefit,3)}, with minor rounding difference of {a-b$benefit}"))


```


The benefit for the focal sample unit with this hypothetical sample unit is `r round(a,3)`. However, as we only sampled 10 sample units for the hypothetical sample set, we may not have an accurate picture of the benefit of this focal sample unit.

-----------


### E. Repeat multiple times

To get a more useful measure of benefit, we calculate the benefit from this focal sample unit across 100 randomly drawn hypothetical sample sets.

```{r, fig.cap ="Fig 6. Example Sample Unit benefit, calculated 100 times. Mean benefit is shown in the vertical dotted line."}

landcover_ha <- StudyArea_hexes$landcover %>% 
  mutate_at(.vars = vars(contains("LC")),.funs = ~(.*area)) %>% ungroup
att.long <- BASSr::prepare_hab_long(as_tibble(landcover_ha))

sample_hexes100 <- BASSr::draw_random_samples(att_cleaned = as_tibble(landcover_ha), 
                                           att.sf = st_centroid(landcover_ha), 
                                           num_runs = 100, nsamples = 10) 

multi_ex <- BASSr::calculate_benefit(grts_res = sample_hexes100, 
                                     HexID = SampleUnitID,
                                     att_long = att.long[att.long$SampleUnitID==exampleHex$SampleUnitID,], 
                                     q=T )

multi_exbroken <- map_df(1:100, 
                         ~quick_ben(d = exampleHex %>% as_tibble %>%
                                      dplyr::select(-geometry) %>%
                                      mutate_at(.vars = vars(contains("LC")),
                                                .funs = ~(.*hexsize)), 
          samples = sample_hexes100$grts_random_sample %>% 
            filter(run == .x) %>% 
             summarize_at(.vars = vars(contains("LC")),.funs = ~sum(.)), 
          land_cover_summary = SA_sum %>% mutate(ha = pHab_SA*study_area_size), 
          col_ = SampleUnitID, pd=F) )
  

ggplot(multi_exbroken, aes(benefit)) +
  geom_density(fill= 'grey') + 
  geom_vline(xintercept = multi_ex$benefit, linetype =2) +
  theme_minimal() +
  labs(x= "Benefit of Example Sample Unit", y = "Density") +
  xlim(0, 0.15)
  
```


The there is quite a large variation in the benefit calculated across all sample set iterations. Our final measure of benefit for the focal sample unit is `r round(multi_ex$benefit,3)`. Now, using the same hypothetical sample sets from this focal sample unit we will calculate the benefit for the other sample units in the study area.

----


### F. Run benefit calculate for all sample units


The total benefit for the example sample unit based on national land cover is shown above (`r  round(multi_ex$benefit,3)`). How does this compare with other values in the study area? We can run the same process on all the other hexagons to find out.

```{r all-benefits, fig.lab = "Fig 7. Single sample set used to calculate benefit estimates across all sample units in a study area"}

one_sample <- quick_ben(
  d = StudyArea_hexes$landcover %>% as_tibble %>% 
             mutate_at(.vars = vars(contains("LC")),.funs = ~(.*hexsize)), 
          samples = sample_hexes$grts_random_sample %>% 
             summarize_at(.vars = vars(contains("LC")),.funs = ~sum(.*hexsize)), 
          land_cover_summary = SA_sum %>% mutate(ha = pHab_SA*study_area_size), 
          col_ = SampleUnitID, pd=F)

ggplot(one_sample, aes(benefit)) + 
  geom_density(fill= 'grey') + 
  geom_vline(xintercept = b$benefit, linetype =2) +
  labs(x= "Benefit", y = "Density") +
  theme_minimal() +
  xlim(0, 0.15)

```

There is as much variation across sample units as there is within iterations of the example sample units. The example sample unit looks to be in the middle of a possibly bimodal distribution. However this is just one hypothetical sample unit. 


The goal is to run this multiple times with many random  hypothetical sample sets. The smaller the number of sample units in a given hypothetical sample set, the wider the distribution in benefits are likely to be across all sample units in a study area and the more variance between runs. So the size of the sample set should be small and iterations nunmerous. 

I ran 200 iterations here with sample sets of 10 study unit hexagons. A key point is to start with hectares (or another measure of area) of land cover rather than percent or proportion or the benefit calculation will have incorrect estimates when the focal sample unit is added to the hypothetical sample set.


```{r 200iterations-1, fig.cap="Fig 7. Mean benefit distribution across 200 iterations of the benefit calculation. The mean benefit for the example hexagon is shown in the dotted line"}

set.seed(1234)
sample_hexes2 <- BASSr::draw_random_samples(att_cleaned = as_tibble(landcover_ha), 
                                           att.sf = st_centroid(landcover_ha), 
                                           num_runs = 200, nsamples = 10) 

benefits2 <- calculate_benefit(sample_hexes2, HexID = SampleUnitID,att_long =  att.long, quick = T, 
                               output = 'mean.benefit')

ggplot(benefits2, aes(benefit)) + 
  geom_density(fill= 'grey') + 
  geom_vline(xintercept = benefits2$benefit[benefits2$SampleUnitID==exampleHex$SampleUnitID], 
             linetype =2) +
  labs(x= "Benefit", y = "Density") +
  theme_minimal() +xlim(0, 0.15)


```

Running multiple iterations narrows the benefit distribution quite a bit.Lets have a look it what a larger sample set size will look like. I next ran 200 interations with 20 samples in each hypothetical sample unit.


```{r 200iterations-2, fig.cap="Fig 8. Mean benefit distribution across 200 iterations of the benefit calculation. The mean benefit for the example hexagon is shown in the dotted line. 20 samples are drawn per iteration"}
set.seed(1234)
sample_hexes3 <- BASSr::draw_random_samples(att_cleaned = as_tibble(landcover_ha), 
                                           att.sf = st_centroid(landcover_ha), 
                                           num_runs = 200, nsamples = 20) 

benefits3 <- calculate_benefit(sample_hexes3, HexID = SampleUnitID,att_long =  att.long, quick = T, 
                               output = 'mean.benefit')

ggplot(benefits3, aes(benefit)) + 
  geom_density(fill= 'grey') + 
  geom_vline(xintercept = benefits3$benefit[benefits3$SampleUnitID==exampleHex$SampleUnitID], 
             linetype =2) +
  labs(x= "Benefit", y = "Density") +
  theme_minimal() +xlim(0, 0.15)


```

This shifts the benefits for any given sample unit smaller, as the benefit of adding an individual sample unit to the sample set is smaller, but the distribution remains roughly the same and the example study unit hexagon in the same position.

## Represent the Study Area
After running multiple iterations of drawing hypothetical sample sets we can then show how the benefit of adding a sample unit to the draw is expected to increase representivity of the final surveyed sample units.

```{r, fig.width=4, fig.height=4, fig.align='center', fig.cap = "Fig 9. The plotted mean benefit estimate running 200 iterations of hypothetical sample sets each of 10 sample units." }
benefit_sf <- 
ggplot(benefits2 %>% right_join(landcover_ha) %>% st_as_sf) +
  geom_sf(aes(fill = benefit)) + scale_fill_viridis_c() + theme_linedraw() +
   ggspatial::annotation_scale(location = "bl", width_hint = 0.4) 
benefit_sf
```


---------------
---------------

## Appendix - Analysis of Benefit Calculation

How does this arise? If we examine the difference between a sample unit's percentage of habitat and that percent found in the study area, we see that on average

```{r, fig.width=4, fig.height=4, fig.cap = "Fig 10. Smoothed curves of how the difference between a sample unit and study area land cover percentage affects the mean benefit for that sample unit. Black line shows the smoothed curve across all land cover types, while the grey line shows the individual land cover types. Only land cover types with greater than 1\\% coverage of study area are show individually here."}
benefits2 %>% right_join(hex_LC) %>% 
  # filter(pHab_SA > 0.05) %>% 
ggplot(
  aes(#round((
    pHab - pHab_SA*100,
    # )5,1)/5, 
benefit )) +
  geom_point(alpha = 0.2, colour = 'grey')+
  geom_smooth(
    data = . %>% filter(pHab_SA > 0.01),
    aes(group = lc), 
              colour= 'darkgrey',
               alpha = 0.5, 
              se=F) +
  geom_smooth(colour = 'black') +
  theme_minimal() +
  labs(x = "Difference  in % between Sample Unit and Study Area", y = "Mean Benefit") 

# + 
#   stat_summary(fun.data = 'mean_cl_boot', geom = 'ribbon', fill = 'grey', na.rm=T, alpha = 0.5) +
#   stat_summary(fun.y = 'mean', geom = 'line') +
#   
#   theme_minimal()
```


In this study area, for many of the sample units that are very different from the study area have higher benefits. This appears to be mostly driven by the mean absolute difference rather than the maximum (see the next two figures for visualizations).



```{r, fig.cap=" Fig 11. Maximum absolute difference across all land cover classes between sample unit and study area and how that relelates to benefit estimate for that hexagon."}
benefits2 %>% right_join(hex_LC) %>% 
  group_by(SampleUnitID) %>% 
  slice(which.max(abs(pHab-pHab_SA*100))) %>% 
  # filter(pHab_SA > 0.05) %>% 
ggplot(
  aes(#round((
    abs(pHab - pHab_SA*100),
    # )5,1)/5, 
benefit )) +
  geom_point(alpha = 0.2, colour = 'grey')+
 
  geom_smooth(colour = 'black') +
  theme_minimal() +
  labs(x = "Difference  in % between Sample Unit and Study Area", y = "Mean Benefit") 
```



```{r, fig.cap="Fig. 12. Mean absolute difference across all land cover classes between sample unit and study area and how that relelates to benefit estimate for that hexagon."}
ben_diffs <- benefits2 %>% right_join(hex_LC) %>% 
  group_by(SampleUnitID, benefit) %>% 
  summarize(phabD = mean(abs(pHab-pHab_SA*100))) %>% 
  ungroup #%>% 
  # filter(pHab_SA > 0.05) %>% 
ggplot(ben_diffs,
  aes(
    phabD,benefit )) +
  geom_point(alpha = 0.2, colour = 'grey')+
  geom_smooth(colour = 'black') +
  theme_minimal() +
  labs(x = "Mean Absolute Difference in % between Sample Unit and Study Area", y = "Mean Benefit") +
  xlim(0, 13.1)
```

So sample units that are more different than the Study Area on average have a higher benefit. 

```{r}
ARU_locs <- run_grts_on_BASS(200, study_area_results = 
                    left_join(landcover_ha,benefits2) %>% 
                   mutate(inclpr = benefit),hexid = SampleUnitID,
                   nARUs =  30, os = 0, idcol = "StudyAreaID")
```




```{r, fig.cap="Fig. 12. The number of draws a sample unit was selected out of 200 GRTS runs of 30 ARUs based on the benefit (left) and absolute difference from the study area (right)"}



z <-map(1:200,~ARU_locs[[.x]]@data %>% filter(panel == "PanelOne") %>% 
  .[["SampleUnitID"]] %>%  as.character)
z_out <- map_df(1:200,~ARU_locs[[.x]]@data %>% filter(panel == "PanelOne") %>% mutate(r = .x) )

z_200 <- 
z_out %>% group_by(SampleUnitID) %>% 
  summarize(n=n())
diff_ben <- 
left_join(ben_diffs, z_200) %>% 
  replace_na(list(n=0)) %>% 
  ggplot(aes(phabD, n)) + geom_point()+
  labs(x = "Mean absolute percentage difference from study area", y = "Number of times chosen for sampling out of 200")

ben_ben <- left_join(ben_diffs, z_200) %>% 
  replace_na(list(n=0)) %>% 
  ggplot(aes(benefit, n)) + geom_point() +
  labs(x = "Mean benefit", y = "Number of times chosen for sampling out of 200")


ben_ben + diff_ben


```


This does lead to an increased chance of being selected (when ignoring cost), but the effect is fairly small. It does mean that hexagons that closely mimic the study area are less likely to be selected. We can see that even when cost is included, in this study area the chances of being selected do not vary much across all sample units.

```{r}
grts_res <- 
left_join(StudyArea_hexes$landcover, z_200) %>% 
  ggplot() +
  geom_sf(aes(fill = n)) +
  scale_fill_viridis_c()+
  theme_linedraw() +
   ggspatial::annotation_scale(location = "bl", width_hint = 0.4) 

benefit_sf+grts_res

```

And habitat representivity in each draw remains pretty good. 


```{r, fig.width=8, fig.height=12}
diff_withSA_draws <- 
z_out %>% 
  pivot_longer(cols = matches("LC\\d"), values_to = "ha", names_to = "lc") %>% 
  group_by(r, lc) %>% 
  summarize(ha = sum(ha)) %>% 
  group_by(r) %>% 
  mutate(phab = ha / sum(ha)) %>% ungroup %>% 
  left_join(SA_sum, by = "lc")

ggplot(diff_withSA_draws, aes(lcc_fac, phab)) + 
  geom_point(position = position_jitter(width = 0.2), alpha = 0.5) +
  geom_point(aes(y = pHab_SA), colour = 'red', shape = "-", size = 10)+
    theme(axis.text.x = element_text(family = 'mono', 
                                   angle= 75, vjust = 1, hjust = 1))
  
```

Here we see the sum of squared error of habitat surveyed to that across the study area in proportion of study area.

```{r}
errors <- diff_withSA_draws %>% 
  mutate(E = ((phab - pHab_SA))**2,
         ae = abs(phab-pHab_SA)*100) %>% 
  group_by(r) %>% 
  summarize(SSE = sum(E),
            MAE = mean(ae)) 
  ggplot(errors,aes(SSE)) +
  geom_density(fill = 'grey') +
  theme_minimal() +
  labs(x = "Sum of Squared Error from Study Area")
```

This is the mean absolute error of study area percent between each GRTS sample and the study area, averaged across all habitat types.

```{r}
ggplot(errors, aes(MAE))+
    geom_density(fill = 'grey') +
  theme_minimal() +
  labs(x = "Mean Absolute Error from Study Area Percent")
```





