---
title: "Getting Started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r}
library(sf)
library(ggplot2)
library(dplyr)
b <- readr::read_rds("../BASSr - Extra/2022-10-27-Data_for_BASSr.rds")
```


```{r}
ggplot() +
  geom_sf(data = b$PSU$Hexagons) +
  geom_sf(data = b$SSU$Points)

ggplot() +
  geom_sf(data = filter(b$SSU$Points, SW_ID == 108))

ggplot() +
  geom_sf(data = b$PSU$Hexagons) +
  geom_sf(data = b$SSU$Points) +
  geom_sf_label(data = b$SSU$Points, aes(label = COMP_ID)) +
  coord_sf(xlim = c(490000, 520000), ylim = c(12500000, 12600000))

```



## Setup

1. Get spatial data - Hexagonal grid (primary spatial units - PSU)
2. Get land cover characteristics for spatial hexes
3. Get cost data
4. Run `full_BASS_run()`    


### 1. Spatial Data
- Spatial data frame (sf object) with 
    - Columns Cell ID (e.g., `ET_Index`)
    - Columns defining land cover (e.g., `CLC15_1`)
    - Columns defining in dominant land cover if no lake and calculating costs (e.g., `D_CLC15`)
- This must be either POINTS or (MULTI)POLYGON (will be converted to points via `st_centroid()`)

> Land cover characteristics should not be percentages, but should be XXXX?

```{r}
lc <- select(b$PSU$CLC, "ET_Index", starts_with("CLC"), "TOTAL_HA_raw") %>%
  clean_forBass(s = "CLC15",
                f_vec =  .$ET_Index,
                id_col = ET_Index)

hex <- select(b$PSU$Hexagons, "ET_Index") %>%
  left_join(lc, by = "ET_Index") %>%
  select(-"Province") %>%
  st_transform(3395)  # Otherwise wrong locations from run_grts_on_bass()

ggplot(data = hex) +
  geom_sf()
```

### 2. Land cover
- Data frame with columns Cell ID (e.g., `ET_Index`) and ...
    - Columns defining land cover (e.g., CORINE land cover CLC `CLC15_1`)
    - Column defining Stratum (e.g., `Province`)

```{r}
ggplot(data = hex, aes(fill = LC18)) +
  geom_sf() +
  labs(fill = "Open Water") +
  scale_fill_viridis_c()

ggplot(data = hex, aes(fill = LC06)) +
  geom_sf() +
  labs(fill = "Mixed Forest") +
  scale_fill_viridis_c()

b$CLC_guide
```


### Basic Run
```{r}
d <- full_BASS_run(num_runs = 10,
                   nsamples = 3,
                   att.sp =  hex,
                   att = lc,
                   HexID_ = ET_Index,
                   stratumID = Province,
                   cost = NULL,
                   q = TRUE,
                   noCost = TRUE)

hex_d <- left_join(hex, d, by = "ET_Index")
ggplot(data = hex_d, aes(fill = benefit)) +
  geom_sf() +
  labs(fill = "Benefit") +
  scale_fill_viridis_c()
```

### Weights? (e.g., Akimiski_Island.Rmd)


## 3. Costs

- Need number of ARUs which to deploy

```{r, eval=T}
water_hex <- hex %>%
  filter(LC18 / TOTAL_HA_raw > 0.9) %>%
  pull(ET_Index)

c <- b$PSU$cost_parameters %>%
  st_transform(3395) %>%
  select(ET_Index, pr, sr, wr, basecamps, airportdist_km, cabin_dist_to_air,
         AirportStatus) %>%
  mutate(INLAKE = ET_Index %in% water_hex,
         Province = "ON")

costs <-  estimate_cost_study_area(narus = 3, c, pr, sr, wr, 
                                   dist_base_sa = basecamps,
                                   dist_airport_sa = airportdist_km,
                                   dist2airport_base = cabin_dist_to_air, # Or cabin_dist_to_hwyair?
                                   AirportType = AirportStatus,
                                   vars = cost_vars)

ggplot(data = costs) +
  geom_sf(aes(colour = log(RawCost))) + 
  scale_colour_viridis_c("Cost Estimate")+
  theme(legend.position = c(0,0.7), legend.background = element_rect(fill=NA))


d <- full_BASS_run(num_runs = 10,
                   nsamples = 3,
                   att.sp =  hex,
                   att = lc,
                   HexID_ = ET_Index,
                   stratumID = Province,
                   cost = costs,
                   q = TRUE)

d_hex <- left_join(hex, d, by = "ET_Index")

ggplot(d_hex, aes(fill = inclpr)) + 
  geom_sf() +
  geom_sf(data = filter(c, airportdist_km < 5), fill = "red") +
  scale_fill_viridis_c()

ggplot(d_hex, aes(fill = LogCost)) + 
  geom_sf() +
  geom_sf(data = filter(c, airportdist_km < 5), fill = "red") +
  scale_fill_viridis_c()

c2 <- pivot_longer(c, cols = -c("ET_Index", "INLAKE", "Province", "AirportStatus", "geometry"))
ggplot(filter(c2, name %in% c("pr", "sr", "wr")), aes(fill = value)) +
  geom_sf(size = 0.5) +
  facet_wrap(~name) +
  scale_fill_viridis_c()

ggplot(filter(c2, !name %in% c("pr", "sr", "wr")), aes(fill = value)) +
  geom_sf(size = 0.5) +
  facet_wrap(~name) +
  scale_fill_viridis_c()

```


But perhaps one type of land cover is prioritized. This leads to weights...

See....
```{r}
#calculate_inclusion_probs()
```


Can also calculate separately...

```{r}
costs2 <- st_centroid(costs)
costs2 <- costs2 %>%
  st_coordinates() %>%
  as_tibble() %>%
  bind_cols(costs2)

calculate_inclusion_probs(cost = costs2,
                          hexagon_benefits = select(d, ET_Index, benefit),
                          HexID = ET_Index, StratumID = Province)
```


## Selection probabilities
```{r}
#CRS  expected is 3395 (must be in first step)

ggplot() + 
  geom_sf(data = hex) +
  geom_sf(data = st_as_sf(d, coords = c("X", "Y"), crs = 3395))

sel <- run_grts_on_BASS(n_grts_tests = 1, 
                        study_area_results = d, 
                        nARUs = 12, 
                        os = 0.2,   # default?
                        idcol = "Province",
                        hexid = ET_Index)

ggplot(d_hex) +
  geom_sf() +
  geom_sf(data = sel[[1]]$sites_base)

```

## SSUs

```{r}
lc <- select(b$SSU$CLC, "SW_ID", "COMP_ID", starts_with("CLC"), "TOTAL_m2") %>%
  clean_forBass(s = "CLC15",
                f_vec =  .$COMP_ID,
                id_col = COMP_ID)

hex_ssu <- select(b$SSU$Points, "COMP_ID") %>%
  left_join(lc, by = "COMP_ID") %>%
  select(-"Province") %>%
  st_transform(3395)  # Otherwise wrong locations from run_grts_on_bass()

ggplot(data = hex_ssu) +
  geom_sf()

set.seed(1111)

ssu1 <- full_BASS_run(num_runs = 10, 
                      nsamples = 3, 
                      att = lc, 
                      att.sp = hex_ssu, 
                      HexID_ = COMP_ID,
                      stratumID = Province,
                      cost = NULL,
                      noCost = TRUE,
                      q = TRUE)

ssu2 <- map(unique(hex_ssu$SW_ID),
            ~full_BASS_run(num_runs = 10, 
                           nsamples = 3, 
                           att = filter(lc, SW_ID == .x),
                           att.sp = filter(hex_ssu, SW_ID == .x),
                           HexID_ = COMP_ID,
                           stratumID = Province,
                           cost = NULL,
                           noCost = TRUE,
                           q = TRUE))

ssu_sel <- map(ssu2, 
               ~st_coordinates(.x) %>%
               ~run_grts_on_BASS(n_grts_tests = 1, 
                                 study_area_results = .x, 
                                 nARUs = 12, 
                                 os = 0.2,   # default?
                                 idcol = "Province",
                                 hexid = COMP_ID))

```



## Selection probabilities all together
```{r}
#CRS  expected is 3395 (must be in first step)

ggplot() + 
  geom_sf(data = hex) +
  geom_sf(data = st_as_sf(d, coords = c("X", "Y"), crs = 3395))

d <- full_BASS_run(num_runs = 10,
                   nsamples = 3,
                   att.sp =  hex,
                   att = lc,
                   HexID_ = ET_Index,
                   stratumID = Province,
                   cost = costs,
                   q = TRUE)

d <- run_full_BASS_w_selection(study_areas_hab_cost,
                               sample_units_w_hab,
                               Number_of_Study_areas = 12, # nARUs run_grts_on_BASS()
                               Number_of_sample_units,
                               Size_of_HSS,        # nsamples full_BASS_run()
                               Number_of_HSS,      # num_runs full_BASS_run()
                               Weight_of_benefit,  # benefit_weight full_BASS_run()
                               LandCoverType,
                               RemovedLayers_,
                               Area_of_interest,
                               RandomSeed)

ggplot(d_hex) +
  geom_sf() +
  geom_sf(data = sel[[1]]$sites_base)

```
  
## After 

```{r}
run_grts_on_BASS()
# OR
run_full_BASS_w_selection()
# OR
select_ssu() #?

```


```{r}
downweight_selection_pr()
genSSU()


```

Helpers...?
```{r}
create_study_area()
estimate_cost_study_area()
clean_forBass()

```

## Potential workflow


```{r, eval = FALSE}
# Load hex landcover data
hex <- st_read(...) %>%
  clean_forBass(s = "CLC15", hex_id = "ET_Index")

hex_ssu <- st_read(...) %>%
  clean_forBass(s = "CLC15", psu_hex_id = "ET_Index", ssu_hex_id = "COMP_ID")

# Estimate costs
costs <- st_read(...) %>%
  estimate_cost_study_area(narus = ..., ...)

# Run PSU cost/benefit
psu <- full_BASS_run(hex, costs, num_runs = 10, nsamples = 3, hex_id = "ET_Index")

# Run selection on PSU
psu_sub <- run_grts_on_BASS(bass_run = psu, 
                            n_grts_tests = 1, 
                            n_arus = 12, 
                            os = 0.2,   # default?
                            hex_id = ET_Index)

# Run SSU cost/benefit
hex_ssu <- filter(hex_ssu, .data$ET_Index %in% psu_sub$sites_base$ET_Index)
ssu <- full_BASS_run(hex_ssu, costs_ssu, 
                     num_runs = 10, nsamples = 3, hex_id = "COMP_ID",
                     by = "ET_Index", sub = psu_sub$ET_Index)

# Run selection on SSU
ssu_sub <- run_grts_on_BASS(bass_run = psu, 
                            n_grts_tests = 1, 
                            n_arus = 12, 
                            os = 0.2,   # default?
                            hex_id = ET_Index)

summarize_selection(ssu_sub)

```



  
# Acronyms
- GRTS - generalized random tessellation stratified sample
- BMS
