---
title: "Getting Started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r}
library(sf)
library(ggplot2)
library(dplyr)
```


## Setup

1. Get spatial data - Hexagonal grid (primary spatial units - PSU) with land cover characteristics for each hex
2. Get cost data
3. Run `full_BASS_run()`    


### 1. Spatial Data
- Spatial data frame (sf object) with 
    - Columns Cell/Hex ID (e.g., `ET_Index`, `hex_id`)
    - Columns defining land cover (e.g., CORINE land cover CLC `CLC15_1`)
- This must be either POINT or (MULTI)POLYGON (will be converted to points)

> Land cover characteristics should not be percentages, but should be XXXX?


**Clean up hex data**
```{r}
psu_hex <- clean_land_cover(psu_hex_dirty,  pattern = "CLC0013_") %>%
  units::drop_units() # for plotting, get rid of m^2

ggplot(data = psu_hex, aes(fill = LC01)) +
  geom_sf() +
  scale_fill_viridis_c()
```


### Basic Run
```{r}
d <- full_BASS_run(att_sf = psu_hex, 
                   num_runs = 10,
                   n_samples = 3,
                   hex_id = hex_id)

ggplot(data = d, aes(colour = benefit)) +
  geom_sf() +
  labs(colour = "Benefit") +
  scale_colour_viridis_c()

d_hex <- left_join(psu_hex, st_drop_geometry(d), by = "hex_id")

ggplot(data = d_hex, aes(fill = benefit)) +
  geom_sf() +
  labs(fill = "Benefit") +
  scale_fill_viridis_c()
```

### Including Costs

```{r}
d <- full_BASS_run(att_sf = psu_hex, 
                   num_runs = 10,
                   n_samples = 3,
                   costs = psu_costs,
                   hex_id = hex_id)

d_hex <- left_join(psu_hex, st_drop_geometry(d), by = "hex_id")

ggplot(data = d_hex, aes(fill = inclpr)) +
  geom_sf() +
  labs(fill = "Inclusion\nProbability") +
  scale_fill_viridis_c()
```

What if the costs of that highly beneficial points was much higher?

```{r}
which.max(d$benefit)
high_cost <- psu_costs
high_cost$RawCost[26] <- high_cost$RawCost[26] * 100

d <- full_BASS_run(att_sf = psu_hex, 
                   num_runs = 10,
                   n_samples = 3,
                   costs = high_cost,
                   hex_id = hex_id)

d_hex <- left_join(psu_hex, st_drop_geometry(d), by = "hex_id")

ggplot(data = d_hex, aes(fill = inclpr)) +
  geom_sf() +
  labs(fill = "Inclusion\nProbability") +
  scale_fill_viridis_c()

```

Still high inclusion probability, but other points become relatively 'better'.

### Runs by 'hand'

```{r}
psu_hex <- clean_land_cover(psu_hex_dirty, pattern = "CLC0013_")
samples <- draw_random_samples(psu_hex, num_runs = 10, n_samples = 3)
benefit <- calculate_benefit(psu_hex, samples, hex_id = hex_id)
inc_prob <- calculate_inclusion_probs(benefit, 
                                      costs = psu_costs, 
                                      hex_id = hex_id)

d_hex <- left_join(psu_hex, st_drop_geometry(inc_prob), by = "hex_id")

ggplot(data = d_hex, aes(fill = inclpr)) +
  geom_sf() +
  labs(fill = "Inclusion\nProbability") +
  scale_fill_viridis_c()
```

Alternative pipe

```{r}
final <- psu_hex_dirty %>%
  clean_land_cover(pattern = "CLC0013_") %>%
  draw_random_samples(num_runs = 10, n_samples = 3) %>%
  calculate_benefit(psu_hex, samples = ., hex_id = hex_id) %>%
  calculate_inclusion_probs(costs = psu_costs, hex_id = hex_id)
```


### Weights? (e.g., Akimiski_Island.Rmd)


## Calculating Costs

- Need number of ARUs which to deploy

...


## Selection probabilities
```{r}
#CRS  expected is 3395 (must be in first step)

ggplot() + 
  geom_sf(data = psu_hex) +
  geom_sf(data = st_as_sf(d, coords = c("X", "Y"), crs = 3395))

sel <- run_grts_on_BASS(n_grts_tests = 1, 
                        study_area_results = final, 
                        nARUs = 12, 
                        os = 0.2,   # default?
                        hex_id = hex_id)

ggplot(d_hex) +
  geom_sf() +
  geom_sf(data = sel[[1]]$sites_base)

```

## SSUs

```{r}
lc <- select(b$SSU$CLC, "SW_ID", "COMP_ID", starts_with("CLC"), "TOTAL_m2") %>%
  clean_forBass(s = "CLC15",
                f_vec =  .$COMP_ID,
                id_col = COMP_ID)

hex_ssu <- select(b$SSU$Points, "COMP_ID") %>%
  left_join(lc, by = "COMP_ID") %>%
  select(-"Province") %>%
  st_transform(3395)  # Otherwise wrong locations from run_grts_on_bass()

ggplot(data = hex_ssu) +
  geom_sf()

set.seed(1111)

ssu1 <- full_BASS_run(num_runs = 10, 
                      nsamples = 3, 
                      att = lc, 
                      att.sp = hex_ssu, 
                      HexID_ = COMP_ID,
                      stratumID = Province,
                      cost = NULL,
                      noCost = TRUE,
                      q = TRUE)

# Seletion - run_grts_on_BASS()

ssu2 <- map(unique(hex_ssu$SW_ID),
            ~full_BASS_run(num_runs = 10, 
                           nsamples = 3, 
                           att = filter(lc, SW_ID == .x),
                           att.sp = filter(hex_ssu, SW_ID == .x),
                           HexID_ = COMP_ID,
                           stratumID = Province,
                           cost = NULL,
                           noCost = TRUE,
                           q = TRUE))

ssu_sel <- map(ssu2, 
               ~st_coordinates(.x) %>%
               ~run_grts_on_BASS(n_grts_tests = 1, 
                                 study_area_results = .x, 
                                 nARUs = 12, 
                                 os = 0.2,   # default?
                                 idcol = "Province",
                                 hexid = COMP_ID))

```



## Selection probabilities all together
```{r}
#CRS  expected is 3395 (must be in first step)

ggplot() + 
  geom_sf(data = hex) +
  geom_sf(data = st_as_sf(d, coords = c("X", "Y"), crs = 3395))

d <- full_BASS_run(num_runs = 10,
                   nsamples = 3,
                   att.sp =  hex,
                   att = lc,
                   HexID_ = ET_Index,
                   stratumID = Province,
                   cost = costs,
                   q = TRUE)

d <- run_full_BASS_w_selection(study_areas_hab_cost,
                               sample_units_w_hab,
                               Number_of_Study_areas = 12, # nARUs run_grts_on_BASS()
                               Number_of_sample_units,
                               Size_of_HSS,        # nsamples full_BASS_run()
                               Number_of_HSS,      # num_runs full_BASS_run()
                               Weight_of_benefit,  # benefit_weight full_BASS_run()
                               LandCoverType,
                               RemovedLayers_,
                               Area_of_interest,
                               RandomSeed)

ggplot(d_hex) +
  geom_sf() +
  geom_sf(data = sel[[1]]$sites_base)

```
  
## After 

```{r}
run_grts_on_BASS()
# OR
run_full_BASS_w_selection()
# OR
select_ssu() #?

```


```{r}
downweight_selection_pr()
genSSU()


```

Helpers...?
```{r}
create_study_area()
estimate_cost_study_area()
clean_forBass()

```

## Potential workflow


```{r, eval = FALSE}
# Load hex landcover data
hex <- st_read(...) %>%
  clean_forBass(s = "CLC15", hex_id = "ET_Index")

hex_ssu <- st_read(...) %>%
  clean_forBass(s = "CLC15", psu_hex_id = "ET_Index", ssu_hex_id = "COMP_ID")

# Estimate costs
costs <- st_read(...) %>%
  estimate_cost_study_area(narus = ..., ...)

# Run PSU cost/benefit
psu <- full_BASS_run(hex, costs, num_runs = 10, nsamples = 3, hex_id = "ET_Index")

# Run selection on PSU
psu_sub <- run_grts_on_BASS(bass_run = psu, 
                            n_grts_tests = 1, 
                            n_arus = 12, 
                            os = 0.2,   # default?
                            hex_id = ET_Index)

# Run SSU cost/benefit
hex_ssu <- filter(hex_ssu, .data$ET_Index %in% psu_sub$sites_base$ET_Index)
ssu <- full_BASS_run(hex_ssu, costs_ssu, 
                     num_runs = 10, nsamples = 3, hex_id = "COMP_ID",
                     by = "ET_Index", sub = psu_sub$ET_Index)

# Run selection on SSU
ssu_sub <- run_grts_on_BASS(bass_run = psu, 
                            n_grts_tests = 1, 
                            n_arus = 12, 
                            os = 0.2,   # default?
                            hex_id = ET_Index)

summarize_selection(ssu_sub)

```



  
# Acronyms
- GRTS - generalized random tessellation stratified sample
- BMS
