---
title: "Getting Started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r}
library(BASSr)
library(sf)
library(ggplot2)
library(dplyr)
```


## Setup

1. Get spatial data - Hexagonal grid (primary spatial units - PSU) with land cover characteristics for each hex
2. Get cost data
3. Run `full_BASS_run()`    


### 1. Spatial Data
- Spatial data frame (sf object) with 
    - Columns Cell/Hex ID (e.g., `ET_Index`, `hex_id`)
    - Columns defining land cover (e.g., CORINE land cover CLC `CLC15_1`)
- This must be either POINT or (MULTI)POLYGON (will be converted to points)

> Land cover characteristics should not be percentages, but should be XXXX?


**Clean up hex data**
```{r}
psu_hex <- clean_land_cover(psu_hex_dirty,  pattern = "CLC0013_") %>%
  units::drop_units() # for plotting, get rid of m^2

ggplot(data = psu_hex, aes(fill = LC01)) +
  geom_sf() +
  scale_fill_viridis_c()
```


### Basic Run
```{r}
d <- full_BASS_run(att_sf = psu_hex, 
                   num_runs = 10,
                   n_samples = 3,
                   hex_id = hex_id)

ggplot(data = d, aes(colour = benefit)) +
  geom_sf() +
  labs(colour = "Benefit") +
  scale_colour_viridis_c()

d_hex <- left_join(psu_hex, st_drop_geometry(d), by = "hex_id")

ggplot(data = d_hex, aes(fill = benefit)) +
  geom_sf() +
  labs(fill = "Benefit") +
  scale_fill_viridis_c()
```

### Including Costs

```{r}
d <- full_BASS_run(att_sf = psu_hex, 
                   num_runs = 10,
                   n_samples = 3,
                   costs = psu_costs,
                   hex_id = hex_id)

d_hex <- left_join(psu_hex, st_drop_geometry(d), by = "hex_id")

ggplot(data = d_hex, aes(fill = inclpr)) +
  geom_sf() +
  labs(fill = "Inclusion\nProbability") +
  scale_fill_viridis_c()
```

What if the costs of that highly beneficial points was much higher?

```{r}
which.max(d$benefit)
high_cost <- psu_costs
high_cost$RawCost[26] <- high_cost$RawCost[26] * 100

d <- full_BASS_run(att_sf = psu_hex, 
                   num_runs = 10,
                   n_samples = 3,
                   costs = high_cost,
                   hex_id = hex_id)

d_hex <- left_join(psu_hex, st_drop_geometry(d), by = "hex_id")

ggplot(data = d_hex, aes(fill = inclpr)) +
  geom_sf() +
  labs(fill = "Inclusion\nProbability") +
  scale_fill_viridis_c()

```

Still high inclusion probability, but other points become relatively 'better'.

### Runs by 'hand'

```{r}
psu_hex <- clean_land_cover(psu_hex_dirty, pattern = "CLC0013_")
samples <- draw_random_samples(psu_hex, num_runs = 10, n_samples = 3)
benefit <- calculate_benefit(psu_hex, samples, hex_id = hex_id)
inc_prob <- calculate_inclusion_probs(benefit, 
                                      costs = psu_costs, 
                                      hex_id = hex_id)

d_hex <- left_join(psu_hex, st_drop_geometry(inc_prob), by = "hex_id")

ggplot(data = d_hex, aes(fill = inclpr)) +
  geom_sf() +
  labs(fill = "Inclusion\nProbability") +
  scale_fill_viridis_c()
```

Alternative pipe

```{r}
final <- psu_hex_dirty %>%
  clean_land_cover(pattern = "CLC0013_") %>%
  draw_random_samples(num_runs = 10, n_samples = 3) %>%
  calculate_benefit(psu_hex, samples = ., hex_id = hex_id) %>%
  calculate_inclusion_probs(costs = psu_costs, hex_id = hex_id)
```


### Weights? (e.g., Akimiski_Island.Rmd)


## Calculating Costs

- Need number of ARUs which to deploy

...


## Selection probabilities
```{r}

ggplot() + 
  geom_sf(data = psu_hex) +
  geom_sf(data = st_as_sf(final, coords = c("X", "Y")))

sel <- run_grts_on_BASS(probs = final, 
                        num_runs = 1, 
                        nARUs = 12, 
                        os = 0.2)

ggplot(d_hex) +
  geom_sf() +
  geom_sf(data = sel[[1]]$sites_base)

```

## SSUs

...


