---
title: "Speed bass testing"
author: "David Hope"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
vignette: >
  %\VignetteIndexEntry{Speed bass testing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup}

library(BASSr)

?speedbass
```


The goal here is to explore the speed bass funtion to make sure it is working as expected. I'll start by creating some data


```{r}

n_land_covers <- 15
study_area_size <- 1E6
samplesize <- 10

hex_equal <- rep(study_area_size/1000/n_land_covers,n_land_covers)

samples_equal <-  rep(study_area_size/1000*samplesize/n_land_covers, n_land_covers) 

total_equal <- rep(study_area_size/n_land_covers, n_land_covers)
cat("sample plus hex \t total \t sample\t difference\t desired direction\tincluded\n")
ex1 <- speedbass(hex = hex_equal, sample = samples_equal, total = total_equal, printDets = T)


```

Explore this more...
```{r}
sprintf(" The land cover percent for the sample will be %s", 
        study_area_size/1000*samplesize/n_land_covers / sum(samples_equal) )

sprintf(" The land cover percent for all land cover types in the hexagon will be %s", 
        study_area_size/1000/n_land_covers / sum(hex_equal) )

sprintf(" The land cover percent for the total study area for each land cover type will be %s", 
        study_area_size/n_land_covers / sum(total_equal) )


```

Therefore if they are equal across land cover types in the hexagon, study area, and sample, there is no benefit to adding a value that continues to maintain this equality between sample and total study area. Leading to a total benefit of `r ex1`.

However if the hexagon does improve the percentage, we should see a benefit


```{r}

samples_unequal <- c(0,rep(study_area_size/1000*(samplesize-1)/n_land_covers, n_land_covers-1) )
cat("sample plus hex \t total \t sample\t difference\t desired direction\tincluded\n")
ex2 <- speedbass(hex = hex_equal, sample = samples_unequal, total = total_equal, printDets = T)

```

Lets increase this all the way to every land cover being a bad fit in the sample

```{r}
samples_random <- sample(1:3000, n_land_covers, replace = T)
ss <- samples_random  / sum(samples_random) * sum(samples_equal) 

cat("sample plus hex \t total \t sample\t difference\t desired direction\tincluded\n")
ex3 <- speedbass(hex = hex_equal, sample = ss, total = total_equal, printDets = T)


```

Now the total benefit to that is  `r ex3`, which is less than benefit for the second example because all the samples have quite a bit of 


What if we have a homogeneous habitat? The obvious answer is that there would be zero benefit to any given hexagon.
Lets check to make sure.

```{r}
n_land_covers <- 1

hex_equal <- rep(study_area_size/1000/n_land_covers,n_land_covers)

samples_equal <-  rep(study_area_size/1000*samplesize/n_land_covers, n_land_covers) 

total_equal <- rep(study_area_size/n_land_covers, n_land_covers)
cat("sample plus hex \t total \t sample\t difference\t desired direction\tincluded\n")
ex4 <- speedbass(hex = hex_equal, sample = samples_equal, total = total_equal, printDets = T)

ex4

```

A given hexagon will have zero benefit because as above in example 1, the random sample is already perfectly matching the 
study area.
